|

### 2021-12-09 04-Chp-Systematic-Review-Results.Rmd
Range of small fixes (randomised, ApoE, etc)

@@

Following the methods outlined in the previous chapter, here I present the results of the comprehensive systematic review. The use of the evidence identified by the review in future chapters is discussed. Finally, I make recommendations for future methodological work around the inclusion of preprints and Mendelian randomisation studies in evidence synthesis exercises.

@@

The database search identified `r num_to_text(23447)` records, of which `r comma(7338)` were duplicated records. Of the random sample of 500 records screened to ensure the accuracy of the study design filters, no eligible records were identified. As expected, many of those excluded by the filters were commentaries/educational articles, or described basic science studies. <!-- TODO include some example citations here -->

@@

Following de-duplication, the titles and abstracts of `r num_to_text(16109)` records were assessed for eligibility. `r num_to_text(387)` were deemed potentially eligible, and the full text records for these were accessed and screened.

The PRISMA flow diagram <!-- TODO CITATION NEEDED - cite PRIMSA flow paper here, if published --> presented in Figure \@ref(fig:prisma-flow-fig) illustrates the movement of articles through the review. To highlight the contribution of preprint archives to the review, the flow diagram delineates between those records captured through databases searches (presented on the left of the diagram) and those captured by the search tool described previously (presented in grey on the right of the diagram).

Common reasons for exclusion at the full text-stage included studies reporting on ineligible exposures (N = `r n_wrong_exposure` ; most commonly an ineligible lipid fraction) or outcomes (N = `r n_wrong_outcome`; e.g. change in cognitive scores), or usin an ineligible study design (N = `r n_study_design`).

@@

For the assessment of the intra-/inter-rater reliability, the estimated values of Gwet's $AC1$ were interpreted against the categories presented in Table \@ref(tab:gwet-table). For the inter-rater reliability, agreement was "almost perfect" ($AC1$ = `r agreeInter_coeff[1]`, $kappa$ = `r agreeInter_coeff[2]`, Table \@ref(tab:agreeInter-table)). Similarly for intra-rater reliability, agreement was "almost perfect" ($AC1$ = `r agreeIntra_coeff[1]`, $kappa$ = `r agreeIntra_coeff[2]`, Table \@ref(tab:agreeIntra-table)). The discrepancy between the $AC1$ and $kappa$ coefficients illustrates the sensitivity of $kappa$ to imbalanced marginals, caused in this sample by a large imbalance  towards exclusion.[@feinstein1990]

@@

The majority of included studies described non-randomised analyses, with the only two included randomised controlled trials (the Heart Protection Study/British Heart Foundation trial[@heartprotectionstudycollaborativegroup2002] and the JUPITER trial[@ridker2008]) both reporting on the effect of statin use on incidence of all-cause dementia in older adults. A similarly small number of Mendelian randomisation studies were identified, several of which employed a two-sample approach using summary statistics from the same published genome wide association studies (GWAS), leading to complications in the synthesis (see Section \@ref(sys-rev-res-AD)).

@@

As illustrated in Figure \@ref(fig:cohortLocations), the majority of reports described studies conducted in high-income countries, with the most high represented region being North America. Of interest, several of the included studies were conducted in Taiwan (`r n_taiwan`), all but one of which made use of the Taiwan National Health Insurance database.

@@

In the following sections, analyses are grouped first by outcome (all-cause dementia, Alzheimer's disease and vascular dementia) and then are further stratified by exposure category.

@@

(ref:statinsRCT-cap) Random-effects meta-analysis of randomised controlled trials examining statin statins on all-cause dementia. D1-D5 and O refer to the five risk-of-bias domains and single overall judgement in the RoB2 tool, while the risk of bias due to missing evidence is shown beside the overall summary diamond in each forest plot presented above.

@@

Finally, a single Mendelian randomisation analysis was identified examining the effect of  lowered LDL-c levels on the risk of all-cause dementia via genetic inhibition of the 3-hydroxy-3-methylglutaryl-coenzyme A reductase (HMGCR), emulating statin treatment (see Section \@ref(intro-statins) for more details of the statin mechanism of action). This analysis provided weak evidence for an effect (`r mrStatin$Dementia$estimate`).

@@

@@

Two Mendelian randomisation studies looked at lipid-lowering specifically as a result of HMGCR inhibition, mediated by SNPs in the gene (rs172338484 and rs12916).`r mrStatin$AD$citations` The first used a one sample approach (SNP-exposure and SNP-outcome associations are estimated using the same dataset) in a large Copenhagen-based cohort, while the second made use of summary level data obtained from the Global Lipids Genetic Consortium (SNP-exposure) and the International Genomics of Alzheimer's Project (SNP-outcome). Meta-analysis of these estimates provided weak evidence of an effect (`r mrStatin$AD$estimate`, Figure \@ref(fig:mrStatinADFig)).

@@

Finally, there were several identified Mendelian randomisation studies examining the effect of genetically lowered LDL-c on Alzheimer's disease risk (Figure \@ref(fig:mrDuplication)). However, all of these studies took a two-sample approach, making use of summary statistics from the GLGC and IGAP consortia. Due to this overlap, which would result in a falsely precise estimate caused by multiple counting of the same participants, no meta-analysis of these studies was performed. However, across each analysis using varying number of SNPS, no evidence for an effect was observed (Figure \@ref(fig:mrDuplication)).

@@

@@

A single Mendelian randomisation analysis was identified that examining the effect of genetically lowered LDL-c levels via HMGCR inhibition on the risk of vascular dementia, which provided some evidence for an effect (`r mrStatin$VaD$estimate`).

@@

Few studies investigated the effect of individual lipid fractions. Only for total cholesterol was there greater than one result reported (Figure \@ref(fig:lipidFractionsVaD)), and a meta-analysis of these found weak evidence for an effect (`r n_effect(obsLipids$VaD_TC)`). A single study provided evidence on the other three fractions,[@yoshitake1995] and similar found minimal evidence of an effect LDL-c (`r n_effect(obsLipids$VaD_LDL,F)`), HDL-c (`r n_effect(obsLipids$VaD_HDL,F)`) or triglycerides (`r n_effect(obsLipids$VaD_TG,F)`).

@@

@@

As shown in the forest plots in the above sections, the risk-of-bias assessments are presented alongside their corresponding numerical result. A more detailed discussion of the sources and directions of bias in each result is presented in Chapter \@ref(tri-heading), and so this section presents a brief summary of the biases observed in each study design.

@@

Following best practice, any result judged to be at critical risk of bias should be excluded from any quantitative analyses.[@sterne2016] `r num_to_text(n_critical,T)` observational studies were excluded on this basis, predominantly due to a lack of adjustment for any potentially important confounders (i.e. the study reported unadjusted estimates).`r critical_citations`

@@

There were `r n_dr_initial` studies initially considered eligible for the dose-response meta-analysis, as they appear to provide data on risk of dementia outcomes across several categories of lipid exposure. However, following data extraction, `r num_to_text(n_dr_excluded)` were excluded as they did not reported the relevant information needed (most commonly, the cut-off measures for each category).

@@

@@

<!-- &nbsp; -->

<!----------------------------------------------------------------------------->
<!-- (ref:preprintWeights-caption) Amount of information added by inclusion of preprinted results, estimated from the weight assigned to the study in a random-effect meta-analysis  -->

<!-- (ref:preprintWeights-scaption) Added information due to inclusion of preprints -->

<!-- ```{r preprintWeights-table, message=FALSE, results="asis", echo = FALSE} -->
<!-- ``` -->
<!----------------------------------------------------------------------------->

<!-- TODO Add extra column with details of Figure showing met-analysis -->

<!-- &nbsp; -->

Investigation of the publication status of the two preprints reporting studies not also identified by the main search found that, allowing for a two-year lag, only one had been subsequently published.[@andrews2019; @andrews2021]

@@

There was some indication of a protective effect of statins on all-cause and Alzheimer's disease dementia when looking at solely at observational studies. This finding was not supported by evidence from the two available RCTs, or by studies that emulated statin treatment using a genetic proxy, suggesting that these findings may be a result of differing exposure windows (e.g. mid-life lipid lowering in non-randomised studies versus life-course lipid exposure in Mendelian randomisation studies and late-life lipid reduction in RCTs) or alternatively due to biases within the non-randomised studies.

@@

Across dementia outcomes, the majority of studies were non-randomised studies of lipids or lipid regulating agents. This distribution of evidence between analytical designs is to be expected. Randomised controlled trials of dementia are particularly challenging, as the follow-up made necessary by the long latent period of the condition makes trials logistically difficult and financial expensive. Similarly, Mendelian randomisation is a comparatively new study design, and studies employing it in relation to dementia outcomes only appear in the evidence base in recent years, as illustrated in Figure \@ref(fig:typeByYear). This recent increase is likely driven by the availability of summary genome wide association studies (GWAS) that form the basis of the two-sample Mendelian randomisation approach.

@@

Risk of bias across the individual results was generally high. The causes and expected impact of these biases on the results are discussed in more detail in Chapter \@ref(tri-heading). Of particular interest to this chapter, however, was the high risk of bias due to missing evidence observed for observational studies of lipid levels. In many cases, estimates are known to be missing not at random from meta-analyses due to preferential reporting of significant results, leading to high risk of bias due to missing evidence. These missing estimates were most commonly identified via analysis of conference abstract/final publication pairs.[@yamada2009conf; @yamada2009] In addition, some authors stated outright that non-significant results were not reported (e.g. "The other lipid variables not significantly associated with dementia and Alzheimer’s disease ... were not reported in the table."[@ancelin2013]). However, as all identified missing results are likely to be non-significant, they would not be expected to have a substantial impact on their respective meta-analysis (which provided weak evidence for an effect), other than increasing the precision of the summary estimate.

Finally in terms of generalisability, despite a large proportion of included studies being conducted in the Western world (Figure \@ref(fig:cohortLocations)), the applicability of the results to other populations is aided by the inclusion of several studies which made use of data from the Taiwan health insurance database.

@@

While conducting this review, I identified several previous systematic reviews of this topic.[@chu2018; @yang2020; @muangpaisan2010; @poly2020;@kuzma2018stroke; @kuzma2018risk] However, to my knowledge, this review is the first to use established domain based assessments tools (for example, the RoB 2 tool for randomized controlled trials)[@sterne2019] to assess the risk of bias in included studies. The majority of the highly cited reviews on this topic either do not formally consider risk of bias in the observational studies they include,[@chu2018;@power2015] or used a non-domain-based assessment tool (e.g. the Newcastle-Ottowa Scale).[@poly2020]

I identified one previous review of Mendelian randomisation studies examining risk factors for Alzheimer's disease. However, this review was conducted prior to the majority of Mendelian randomisation studies included in this review being published and extracted results including SNPs in the _ApoE4_ genetic region (see the following section for a discussion of the bias this introduces).

Despite these differences in time scales and methodology, the duplication of work across systematic reviews (including this one) is substantial. In retrospect, an alternative approach to conducting a further systematic review from scratch, known as an umbrella review or review-of-reviews,[@aromataris2015; @smith2011] could have been employed. This study design uses other systematic reviews rather than primary studies as the unit of analysis, and would have enabled more efficient identification of relevant primary studies to which the methods which sets this review apart could have been applied.

@@

Mendelian randomisation is a powerful analytical technique, using natural variation in participants genomes to identify causal links between a genetically determined risk factor and an outcome, given that the three core assumptions detailed in Figure \@ref(fig:mrAssumptions) are valid. These assumptions are namely that: i) the genetic variant associates with the risk factor of interest (relevance assumption); ii) the variant-exposure association has no unmeasured confounders (independence assumption); and the variants affect the outcome only through their effect on the risk factor of interest (exclusion restriction assumption). However, inclusion of Mendelian randomisation as an acceptable study design in this review was complicated by a number of factors.

@@

Firstly, this study design is relatively new, particularly when compared to randomised trials or cohort studies. Figure \@ref(fig:typeByYear) demonstrates that Mendelian randomisation studies only begin to appear in the evidence base much later than NRSE/NRSI, likely due to the limited availability of large scale GWAS datasets needed for two-sample Mendelian randomisation analyses. As such, the process and tools for systematically assessing this study design are not as well developed. A key example of this is the absence of validated search filters for Mendelian randomisation studies. This limitation is further complicated by the varying terminology used to describe the method, particularly in the early years of its application, which led to me including general terms for instrumental variable analyses in my search.

Additionally, there is currently no widely used risk-of-bias assessment tool for Mendelian randomisation studies. While a recent commentary provided a checklist for interpreting Mendelian randomisation studies, this guide includes reporting items in their quality checklist.[@davies2018] Reporting quality is an important consideration for Mendelian randomisation studies, as reflected by the recent release of a STROBE MR reporting guidelines,[@skrivankova2021] but is a separate consideration to bias, as discussed in Section \@ref(risk-of-bias). Similarly, a previous review of Mendelian randomisation studies used the Q-Genie tool which was designed to assess the quality of GWAS included in a meta-analysis.[@sohani2015] While this tool assesses the underlying GWAS used, it does not address the additional methodological considerations of the Mendelian randomisation analysis itself. For this review, I used the best available author-devised tool, sourced from a recent review of systematic reviews of Mendelian randomisation studies.[@spiga2021]

As a further stumbling block, Mendelian randomisation lends itself to the analysis of multiple exposure-outcome comparisons as part of a single study, particularly when using a two-sample summary data design. This is particularly relevant to the consideration of bias due to missing evidence. As an example, through snowballing and other measures, I identified at least one relevant Mendelian randomisation study that had not been captured by the search strategy.[@larsson2017] This study examined lipid fractions as one of many risk factors for Alzheimer's disease, though the search would not have been expected to find it given the absence of any lipid-related keywords in the title or abstract. Studies examining multiple risk factors such as this can introduce bias into a systematic review, as it is commonly only those risk factors with statistically significant result that are reported in the abstract and so are captured by a systematic search. These studies are described as "unknown unknown's" in the context of the RoB-ME tool, and are considered to be particularly challenging (as opposed to an analysis that was insufficiently reported to be included in the statistical analysis, or the "known unknown's").

Useful future work to improve the methodology for inclusion of Mendelian randomisation studies in systematic reviews should involve the development of a validated search filter for this study design.[@waffenschmidt2020; @wagner2020] Alternatively, in better-resourced reviews, a broader search (e.g. "risk factor" AND "dementia" AND "Mendelian randomisation") followed by manual review of studies that examined multiple risk factors would be advisable. This was not feasible in the context of this review, given the large number of records to be screened even when using study design filters (N = `r comma(16109)`). Additionally, the value of methods that support traditional bibliographic database search, such as snowballing (forwards and backwards citation chasing) and communication with relevant topic experts, should not be underestimated. Finally, development of a risk-of-bias assessment tool for this study design by a panel of methodologists and analysts would be of substantial benefit.

One item of particular interest is the attenuation of any effects identified by Mendelian randomisation studies following the adjustment for/exclusion of genetic variation in the _ApoE4_ gene region. As discussed in the introduction (see Section \@ref(intro-basic-science)), an increasing number of _ApoE4_ alleles is a major independent risk factor for Alzheimer's disease and so violates the exclusion restriction criteria of Mendelian randomisation studies (Figure \@ref(fig:mrAssumptions)). In all cases, excluding variants in the _ApoE_ region attenuates the observed effect to the null. A clear example of this is Benn _et al_ (2017), where _ApoE_ variants were not sufficient identified and excluded, leading the published paper to report a protective effect of LDL-c on Alzheimer's (`r estimate(0.83, .75,.92,"RR")`).[@benn2017] Following several rapid responses, the data was re-analysed to exclude a larger area around _ApoE_ which attenuated the finding to the null.[@benn2017comment]

@@

As highlighted in Section \@ref(diverse-sources-preprints), this review explicitly sought to synthesize evidence across different publication statuses (preprinted vs. published). Using the tool described in Chapter \@ref(sys-rev-tools-intro), two preprint servers related to health and biomedical sciences were searched as part of this review. The small number of studies returned by the preprint searches (or the absence of any relevant hits in the medRxiv database - see Figure \@ref(fig:prisma-flow-fig)) is largely due to the timing of the preprint searches. The searches for this review were performed in mid-July 2019, but the medRxiv repository, an offshoot of the Epidemiology and Clinical Trials categories of the bioRxiv preprint server, only registered its first preprint 25th June 2019. As such, at the point it was searched, the medRxiv database contained only a very small number of records (N = `r n_at_search["med"]`; Figure \@ref(fig:preprintGrowth)).

@@

(ref:preprintGrowth-cap) Growth of preprint repositories over time. Given the relative sizes of the preprint repositories at the time the searches for this review were conducted (bioRxiv N = `r n_at_search["bio"]`, medRxiv n = `r n_at_search["med"]`), the number of hits returned by each is expected.

@@

Three relevant preprints from bioRxiv were identified. Of note, all three described a Mendelian randomisation study, potentially indicating that more biologically-focused study designs are over-represented in the bioRxiv repository. The added evidential value of including these preprints was described in Section \@ref(sys-rev-including-preprints-res) and indicated that results available only via preprinted reports can contribute additional evidence to a meta-analysis. However, the scale of this contribution is questionable. Preprinted evidence was incorporated into only single meta-analysis in this review, and inclusion of the preprinted result did not substantially impact the summary effect estimate.

@@

As discussed in Section \@ref(dose-response-results), many primary studies did not report important information required for the dose-response meta-analysis, and so could not be included in the synthesis. This limitation was compounded by the expected low response rate to requests for further information from primary authors. While contacting authors is worthwhile, as it can substantially change the conclusion of a systematic review[@meursingereynders2019] and is not too costly to systematic reviewers,[@cooper2019] a far preferable option is that the authors of primary studies readily deposit all relevant study data at the point of publication.

Based on my experience of extracting data for this review, I co-authored a guidance article to aid primary prevention scientists in preparing and sharing their data so that it can easily be incorporated into a evidence synthesis exercise, using a trial of mindfulness interventions as an case study.[@hennessy2021] A copy of this publication is available in Appendix \@ref(published-papers). In an attempt to apply my own guidance, I have invested a substantial amount of time and effort into making the data obtained by this review openly available to other researchers, via a GitHub repository.

@@

I believe there are several aspects where this review is distinct from those already available in the published literature. While several reviews of this research topic exist,[@chu2018; @yang2020; @muangpaisan2010; @poly2020] the overlap between the list of studies included in each is not complete. As part of this review, I have not only performed a original search of primary literature databases, but have also screened the reference lists of comparable reviews to ensure no relevant study has been omitted.

>>>>

<!-- TODO Consider adding a line to the descriptions of each forest plot regarding the direction - what's on the right, whats on the left -->

@@

Following the methods outlined in the previous chapter, I present the results of the comprehensive systematic review here. The use of the evidence identified by the review in future chapters is discussed. Finally, I make recommendations for future methodological work around the inclusion of preprints and Mendelian randomisation studies in evidence synthesis exercises.

@@

The database search identified `r num_to_text(23447)` records, of which `r comma(7338)` were duplicated records. To ensure the accuracy of the study design filters, a random sample of 500 records removed by the filters were screened, and no eligible records were identified. As expected, many of those excluded by the filters were commentaries/educational articles, or described basic science studies. <!-- TODO include some example citations here -->

@@

Following de-duplication, the titles and abstracts of `r comma(16109)` records were assessed for eligibility. Of these, `r comma(387)` were deemed potentially eligible. The full text records for these records were then accessed and screened.

The PRISMA flow diagram presented in Figure \@ref(fig:prisma-flow-fig) illustrates the movement of articles through the review. To highlight the contribution of preprint archives to the review, the flow diagram delineates between those records captured through databases searches (presented on the left of the diagram) and those captured by the search tool described previously (presented in grey on the right of the diagram).

Common reasons for exclusion at the full-text stage included studies reporting on ineligible exposures (Number of studies (N) = `r n_wrong_exposure` ; most commonly an ineligible lipid fraction) or outcomes (N = `r n_wrong_outcome`; e.g. change in cognitive scores), or the use of an ineligible study design (N = `r n_study_design`).

@@

For the assessment of the intra-/inter-rater reliability, the estimated values of Gwet's agreement coefficinet, $AC1$, were interpreted against the categories presented in Table \@ref(tab:gwet-table). For the inter-rater reliability, agreement was "almost perfect" ($AC1$ = `r agreeInter_coeff[1]`, $kappa$ = `r agreeInter_coeff[2]`, Table \@ref(tab:agreeInter-table)). Similarly for intra-rater reliability, agreement was "almost perfect" ($AC1$ = `r agreeIntra_coeff[1]`, $kappa$ = `r agreeIntra_coeff[2]`, Table \@ref(tab:agreeIntra-table)). The discrepancy between the $AC1$ and $kappa$ coefficients illustrates the sensitivity of $kappa$ to imbalanced marginals, caused in this sample by a large imbalance  towards exclusion.[@feinstein1990]

@@

The majority of included studies described non-randomised analyses, with the only two included randomised controlled trials (the Heart Protection Study/British Heart Foundation trial[@heartprotectionstudycollaborativegroup2002] and the JUPITER trial[@ridker2008]) both reporting on the effect of statin use on incidence of all-cause dementia in older adults. A similarly small number of Mendelian randomisation studies were identified, several of which employed a two-sample approach using summary statistics from the same published genome wide association studies (GWAS), leading to complications in the synthesis of these estimates due to the potential for double-counting (see Section \@ref(sys-rev-res-AD)).

@@

As illustrated in Figure \@ref(fig:cohortLocations), the majority of reports described studies conducted in high-income countries, with the most highly represented region being North America. Of interest, several of the included studies were conducted in Taiwan (`r n_taiwan`), all but one of which made use of the Taiwan National Health Insurance database.

@@

In the following sections, analyses are grouped first by outcome (all-cause dementia, Alzheimer's disease and vascular dementia) and are then further stratified by exposure category (lipids, lipid regulating agents).

@@

@@

@@

(ref:statinsRCT-cap) Random-effects meta-analysis of randomised controlled trials examining statins on all-cause dementia. D1-D5 and O refer to the five risk-of-bias domains and single overall judgement in the RoB2 tool, while the risk of bias due to missing evidence is shown beside the overall summary diamond in each forest plot presented above.

Finally, a single Mendelian randomisation analysis was identified examining the effect of  lowered LDL-c levels on the risk of all-cause dementia via genetic inhibition of the 3-hydroxy-3-methylglutaryl-coenzyme A reductase (HMGCR), emulating statin treatment (see Section \@ref(intro-statins) for more details of the statin mechanism of action). This analysis provided weak evidence of an effect (`r mrStatin$Dementia$estimate`).

@@

@@

Finally, there were several identified Mendelian randomisation studies examining the effect of genetically lowered LDL-c on Alzheimer's disease risk (Figure \@ref(fig:mrDuplication)). However, all of these studies took a two-sample approach, making use of summary statistics from the Global Lipid Genetics Consortium (GLGC) and International Genomics of Alzheimer's Project (IGAP) consortia. Synthesis of these results would would result in a falsely precise estimate caused by multiple counting of the same participants,and as a result, no meta-analysis of these studies was performed.

However, across each analysis using varying number of SNPS, no evidence for an effect was observed (Figure \@ref(fig:mrDuplication)).

@@

Two Mendelian randomisation studies looked at lipid-lowering specifically as a result of HMGCR inhibition, mediated by SNPs in the gene (rs172338484 and rs12916).`r mrStatin$AD$citations` The first used a one sample approach (SNP-exposure and SNP-outcome associations are estimated using the same dataset) in a large Copenhagen-based cohort, while the second made use of summary level data obtained from the GLGC (SNP-exposure) and the IGAP (SNP-outcome). Meta-analysis of these estimates provided weak evidence of an effect (`r mrStatin$AD$estimate`, Figure \@ref(fig:mrStatinADFig)).

@@

@@

Few studies investigated the effect of individual lipid fractions. Only for total cholesterol was there greater than one result reported (Figure \@ref(fig:lipidFractionsVaD)), and a meta-analysis of these found weak evidence for an effect (`r n_effect(obsLipids$VaD_TC)`). A single study provided evidence on the other three fractions,[@yoshitake1995] and similarly found minimal evidence of an effect LDL-c (`r n_effect(obsLipids$VaD_LDL,F)`), HDL-c (`r n_effect(obsLipids$VaD_HDL,F)`) or triglycerides (`r n_effect(obsLipids$VaD_TG,F)`).

@@

A single Mendelian randomisation analysis was identified that examined the effect of genetically lowered LDL-c levels via HMGCR inhibition on the risk of vascular dementia, which provided some evidence for an effect (`r mrStatin$VaD$estimate`).

@@

As shown in the forest plots in the above sections, the risk-of-bias assessments are presented alongside their corresponding numerical result. This section presents only a brief summary of the biases observed in each study design, as s detailed discussion of the sources and directions of bias in each result is presented in Chapter \@ref(tri-heading).

@@

Following best practice, any result judged to be at critical risk of bias should be excluded from any quantitative analysis.[@sterne2016] `r num_to_text(n_critical,T)` observational studies were excluded on this basis, predominantly due to a lack of adjustment for any potentially important confounders (i.e. the study reported unadjusted estimates).`r critical_citations`

@@

There were `r n_dr_initial` studies initially considered eligible for the dose-response meta-analysis, as they appear to provide data on risk of dementia outcomes across several categories of lipid exposure. However, following data extraction, `r num_to_text(n_dr_excluded)` were excluded as they did not report the relevant information needed (most commonly, the cut-off measures for each category).

@@

<!-- TODO Reword the above section to make it clearer -->

@@

For one of the three preprints identified, the published version of the preprint was captured by the main search. Investigation of the publication status of the two unpublished preprints found that, allowing for a two-year lag (i.e. up to July 2021), only one had been subsequently published.[@andrews2019; @andrews2021]

@@

There was some indication of a protective effect of statins on all-cause and Alzheimer's disease dementia when looking solely at observational studies. This finding was not supported by evidence from the two available RCTs, or by studies that emulated statin treatment using a genetic proxy, suggesting that these findings may be a result of differing exposure windows (e.g. mid-life lipid lowering in non-randomised studies versus life-course lipid exposure in Mendelian randomisation studies and late-life lipid reduction in RCTs) or alternatively due to biases within the non-randomised studies.

@@

Across dementia outcomes, the majority of studies were non-randomised studies of lipids or lipid regulating agents. This distribution of evidence between analytical designs is to be expected. Randomised controlled trials of dementia are particularly challenging, as the follow-up made necessary by the long latent period of the condition makes trials logistically difficult and financially expensive. Similarly, Mendelian randomisation is a comparatively new study design, and studies employing it in relation to dementia outcomes only appear in the evidence base in recent years, as illustrated in Figure \@ref(fig:typeByYear). This recent increase is likely driven by the availability of summary genome wide association studies (GWAS) that form the basis of the two-sample Mendelian randomisation approach.

@@

Risk of bias across the individual results was generally high. The causes and expected impact of these biases on the results are discussed in more detail in Chapter \@ref(tri-heading). Of particular interest to this chapter, however, was the high risk of bias due to missing evidence observed for observational studies of lipid levels. In many cases, estimates were known to be missing from meta-analyses due to preferential reporting of significant results, leading to high risk of bias due to missing evidence. These missing estimates were most commonly identified through differing descriptions of the same analysis in conference abstract versus the final publication.[@yamada2009conf; @yamada2009] In addition, some authors stated outright that non-significant results were not reported (e.g. "The other lipid variables not significantly associated with dementia and Alzheimer’s disease ... were not reported in the table."[@ancelin2013]). However, as all identified missing results are likely to be non-significant, they would not be expected to have a substantial impact on their respective meta-analysis (which provided weak evidence of an effect), other than increasing the precision of the summary estimate.

Finally in terms of generalisability, despite a large proportion of the included studies being conducted in the Western world (Figure \@ref(fig:cohortLocations)), the applicability of the results to other populations is aided by the inclusion of several studies which made use of data from the Taiwan health insurance database.

@@

While conducting this review, I identified several previous systematic reviews of this topic.[@chu2018; @yang2020; @muangpaisan2010; @poly2020;@kuzma2018stroke; @kuzma2018risk] However, to my knowledge, this review is the first to use established domain based assessments tools (for example, the RoB 2 tool for randomised controlled trials)[@sterne2019] to assess the risk of bias in included studies. The majority of the highly cited reviews on this topic either do not formally consider risk of bias in the observational studies they include,[@chu2018;@power2015] or used a non-domain-based assessment tool (e.g. the Newcastle-Ottowa Scale).[@poly2020]

I identified one previous review of Mendelian randomisation studies examining risk factors for Alzheimer's disease. However, this review was conducted prior to the majority of Mendelian randomisation studies included in this review being published. In addition this previous review extracted the results of analyses which did not account for SNPs in the _Apo_$\mathcal{E}4$ genetic region (see the following section for a discussion of the bias this introduces).

Despite these differences in time scales and methodology, the duplication of work across systematic reviews (including this one) is substantial. In retrospect, an alternative approach to conducting a further systematic review from scratch, known as an umbrella review or review-of-reviews,[@aromataris2015; @smith2011] could have been employed. This study design uses other systematic reviews rather than primary studies as the unit of analysis. Furthermore, it would have enabled more efficient identification of relevant primary studies to which the methods which sets this review apart could have been applied.

@@

Mendelian randomisation is a powerful analytical technique, using natural variation in participants' genomes to identify causal links between a genetically determined risk factor and an outcome, given that the three core assumptions detailed in Figure \@ref(fig:mrAssumptions) are valid. These assumptions are namely that:

* the genetic variant associates with the risk factor of interest (relevance assumption);
* the variant-exposure association has no unmeasured confounders (independence assumption);
* the variants affect the outcome only through their effect on the risk factor of interest (exclusion restriction assumption).

@@

However, inclusion of Mendelian randomisation as an acceptable study design in this review was complicated by a number of factors. Firstly, this study design is relatively new, particularly when compared to randomised trials or cohort studies. Figure \@ref(fig:typeByYear) demonstrates that Mendelian randomisation studies only begin to appear in the evidence base much later than NRSE/NRSI, likely due to the limited availability of large scale GWAS datasets needed for two-sample Mendelian randomisation analyses. As such, the process and tools for systematically assessing this study design are not as well developed. A key example of this is the absence of validated search filters for Mendelian randomisation studies. This limitation is further complicated by the varying terminology used to describe the method, particularly in the early years of its application, which led to me including general terms for instrumental variable analyses in my search.

Additionally, there is currently no widely used risk-of-bias assessment tool for Mendelian randomisation studies. While a recent commentary provided a checklist for interpreting Mendelian randomisation studies, this guide includes reporting items in their quality checklist.[@davies2018] Reporting quality is an important consideration for Mendelian randomisation studies, as reflected by the recent release of the STROBE MR reporting guidelines,[@skrivankova2021] but it is a separate consideration to bias, as discussed in Section \@ref(risk-of-bias). Similarly, a previous review of Mendelian randomisation studies used the Q-Genie tool which was designed to assess the quality of GWAS included in a meta-analysis.[@sohani2015] While this tool assesses the underlying GWAS used, it does not address the additional methodological considerations of the Mendelian randomisation analysis itself. For this review, I used the best available author-devised tool, sourced from a recent review of systematic reviews of Mendelian randomisation studies.[@spiga2021]

As a further stumbling block, Mendelian randomisation lends itself to the analysis of multiple exposure-outcome comparisons as part of a single study, particularly when using a two-sample summary data design. This is particularly relevant to the consideration of bias due to missing evidence. As an example, through snowballing and other measures, I identified at least one relevant Mendelian randomisation study that had not been captured by the search strategy.[@larsson2017] This study examined lipid fractions as one of many risk factors for Alzheimer's disease, though the search would not have been expected to find it given the absence of any lipid-related keywords in the title or abstract. Studies examining multiple risk factors such as this can introduce bias into a systematic review, as it is commonly only those risk factors with a statistically significant result that are reported in the abstract and so are captured by a systematic search. These studies are described as "unknown unknown's" in the context of the RoB-ME tool, and are considered to be particularly challenging (as opposed to an analysis that was insufficiently reported to be included in the statistical analysis, or the "known unknown's").

Useful future work to improve the methodology for inclusion of Mendelian randomisation studies in systematic reviews should involve the development of a validated search filter for this study design.[@waffenschmidt2020; @wagner2020] Alternatively, in better-resourced reviews, a broader search (e.g. "risk factor" AND "dementia" AND "Mendelian randomisation") followed by manual review of studies that examined multiple risk factors would be advisable. This was not feasible in the context of this review, given the large number of records to be screened even when using study design filters (N = `r comma(16109)`). Additionally, the value of methods that support traditional bibliographic database searches, such as snowballing (forwards and backwards citation chasing) and communication with relevant topic experts, should not be underestimated. Finally, development of a risk-of-bias assessment tool for this study design by a panel of methodologists and analysts would be of substantial benefit.

One item of particular interest is the attenuation of any effects identified by Mendelian randomisation studies following the adjustment for/exclusion of genetic variation in the _Apo_$\mathcal{E}4$ gene region. As discussed in the introduction (see Section \@ref(intro-basic-science)), an increasing number of _Apo_$\mathcal{E}4$ alleles is a major independent risk factor for Alzheimer's disease and so violates the exclusion restriction criteria of Mendelian randomisation studies (Figure \@ref(fig:mrAssumptions)). In all cases, excluding variants in the _Apo_$\mathcal{E}$ region attenuates the observed effect to the null. A clear example of this is Benn _et al_ (2017), where _Apo_$\mathcal{E}$ variants were not sufficiently identified and excluded, leading the published paper to report a protective effect of LDL-c on Alzheimer's (`r estimate(0.83, .75,.92,"RR")`).[@benn2017] Following several rapid responses, the data was re-analysed to exclude a larger area around _Apo_$\mathcal{E}$ which attenuated the finding to the null.[@benn2017comment]

@@

As highlighted in Section \@ref(diverse-sources-preprints), this review explicitly sought to synthesize all available evidence, irrespective of publication status (preprinted vs. published). Using the tool described in Chapter \@ref(sys-rev-tools-intro), two preprint servers related to health and biomedical sciences were searched as part of this review. The small number of studies returned by the preprint searches (or the absence of any relevant hits in the medRxiv database - see Figure \@ref(fig:prisma-flow-fig)) is largely due to the timing of the preprint searches. The searches for this review were performed in mid-July 2019, but the medRxiv repository, an offshoot of the Epidemiology and Clinical Trials categories of the bioRxiv preprint server, only registered its first preprint 25th June 2019. As such, at the point it was searched, the medRxiv database contained only a very small number of records (N = `r n_at_search["med"]`; Figure \@ref(fig:preprintGrowth)).

@@

(ref:preprintGrowth-cap) Growth of preprint repositories over time. Given the relative sizes of the preprint repositories at the time the searches for this review were conducted (bioRxiv N = `r n_at_search["bio"]`, medRxiv N = `r n_at_search["med"]`), the number of hits returned by each is expected.

@@

Three relevant preprints from bioRxiv were identified. Of note, all three described a Mendelian randomisation study, potentially indicating that more biologically-focused study designs are over-represented in the bioRxiv repository. The added evidential value of including these preprints was described in Section \@ref(sys-rev-including-preprints-res) and indicated that results available only via preprinted reports can contribute additional evidence to a meta-analysis. However, the scale of this contribution is questionable. Preprinted evidence was incorporated into only a single meta-analysis in this review, and inclusion of the preprinted result did not substantially impact the summary effect estimate.

@@

As discussed in Section \@ref(dose-response-results), many primary studies did not report important information required for the dose-response meta-analysis, and so could not be included in the synthesis. This limitation was compounded by the expected low response rate to requests for further information from primary authors. Contacting authors is worthwhile, as it can substantially change the conclusion of a systematic review[@meursingereynders2019] and is not too costly to systematic reviewers,[@cooper2019] though a preferable option is that the authors of primary studies readily deposit all relevant study data at the point of publication.

Based on my experience of extracting data for this review, I co-authored a guidance article to aid primary prevention scientists in preparing and sharing their data so that it can easily be incorporated into a evidence synthesis exercise, using a trial of mindfulness interventions as a case study.[@hennessy2021] A copy of this publication is available in Appendix \@ref(published-papers). In an attempt to apply my own guidance, I have invested a substantial amount of time and effort into making the data obtained by this review openly available to other researchers, via a GitHub repository.

@@

There are several aspects where this review is distinct from those already available in the published literature. While several reviews of this research topic exist,[@chu2018; @yang2020; @muangpaisan2010; @poly2020] the overlap between the list of studies included in each is not complete. As part of this review, I have not only performed an original search of primary literature databases, but have also screened the reference lists of comparable reviews to ensure no relevant study has been omitted.
